\newcommand{\talkabstract}[4]%
{%
\newpage%
% \newpage%
\subsection*{#1}%
\subsubsection*{#2}%
#4 \par%
{\em{#3}}%
}

\newcommand{\talkabstractwithoutsub}[3]%
{%
\newpage%
% \newpage%
\subsection*{#1}%
#3 \par%
{\em{#2}}%
}

\talkabstract{Everything but directions.}%
{All the other exciting things you can do with routing.}%
{Dennis Luxen, the lead developer of Open Source Routing Machine, a high-performance routing engine, is an algorithm engineer at heart. He holds a MSc as well as a PhD in computer science. His work focuses on highly scalable route planning mapping services. Dennis is currently working at Mapbox. }%
{The talk explores the benefits of a state-of-the-art routing engine that go beyond the means to get simple driving or walking directions. We explore how 
routing is used as a building block not only to sophisticated data analysis but also in other new and exciting areas. Whether it is applied to calcaluate the walkability of neighborhoods, the matching of riders and drivers for carpooling, or even the monitoring for vandalism in the OpenStreetMap database, 
among a couple of further use cases. And in many cases these questions can even be answered in real-time without any noticeable delay, delivering a superior performance.

With the challenge of handling an ever-growing data set, it is important to use methods that handle data efficiently and can keep up with the data growth and change rates. The talk shows the combination of algorithm engineering, originated in the fundamental research of academia, and the crowd-sourced data 
of OpenStreetMap delivers not only superior features but also a superior user experience.}

\talkabstract{I've bought a car for mapping, now what?}%
{First: glue an "OpenStreetMap" sticker to it}%
{Long time mapper who likes to go outside, editor of SHTOSM, a russian OSM news blog, organizer of some mapping parties and conferences. }%
{The author focuses on field mapping: collecting data on the move, both on passenger seat with a notebook, and behind the wheel, when you cannot spare a second to look away from the road. There will be a history of attempts at mapping as much as possible, sometimes with large-scale projects and examples of what commercial providers can do. Entering data in OpenStreetMap is also a major problem: it has to be easy to be done on large scale. The topic affects mapping on feet and on bicycle, since the concepts developed will affect field mapping in general. Finally, the main question: do you really need a 4x4 jeep for mapping Russia?}

\talkabstract{BikeDistrict - The smart city by bicycle}%
{Crowdsourced tools to support urban cyclists and urban bike planning}%
{\emph{The District} is a company based in Milan specialized in digital services for urban mobility. BikeDistrict it is the the main project (www.bikedistrict.org) available for web and mobile, is an interactive OSM-based map for those urban cyclists riding in "not entirely" bike-friendly cities.}%
{In the context of the optimization of resources to devote to bicycle mobility, the two-years experience of the BikeDistrict project will be presented.
BikeDistrict is primarily a web and mobile application working as a free navigation tool, suggesting the urban bike-friendly paths. 
An intelligent "street rating" system has been developed for the BikeDistrict map interface with the double purpose of:
\begin{itemize}
\item allowing the complete customization of the bike path, according to the specific needs of the user,
\item contemporarily collect and aggregate users "street ratings" to better inform the routing algorithm, and the bike network database, with the final aim of providing the best possible routes for the different categories of urban cyclists. 
\end{itemize}
 
BikeDistrict is then configured as an urban bike mobility monitoring and analysis device. BikeDistrict has the ability to identify critical infrastructural gaps, to evaluate the potential benefits of specific interventions on the bicycle network and in general to support the planning activities by providing real-time-data and analysis related to the infrastructure condition and demand patterns. BikeDistrict is capable of collecting valuable information about the bike mobility demand (e.g 400 000 origin/destination itineraries in the city of Milan during 2013) and regarding the perception of quality of the road infrastructure and the bike facilities, expressed directly by the bike users. }

\talkabstract{Opengeofiction}%
{Using the OSM software for mapping a fictional planet}%
{Thilo Stapff has always been fascinated by maps, real or imaginary. 
He studied Mathematics and works as software developer in Frankfurt. He created
Opengeofiction in 2013. Johannes Bouchain, Hamburg, is interested in
(imaginary) cartography since childhood, Web designer/urban planner and
Opengeofiction co-founder.}%
{Opengeofiction (http://ww.opengeofiction.net) is a collaborative platform for the creation of fictional maps, founded in 2013 using the OSM software stack. In this presentation, we want to describe our motivation for working with fictional maps, our fictional mapmaking history, how and why we created Opengeofiction, and how the project is currently developing. We will also highlight some of the technical challenges we encountered.

Employing the OSM software has taken our fictional mapmaking to a whole new level, not only for practical reasons, but also because it has allowed us to transform a somewhat solitary hobby into a collaborative activity within a growing community. While at a first glance Opengeofiction might look like a miniature version of the Openstreetmap project itself, there are also some key differences. We will explain why we consider the OSM software stack to be generally very well suited for a project like ours, but also discuss the areas where it does not quite fit as well.

We're looking forward to present the Opengeofiction project on the SotM-EU 2014 conference and we hope that we'll give an interesting input about how the OSM software can be used in a rather unconventional way.  And we will show you some far away countries where you’ve never been before.}


\talkabstract{Map rendering beyond Mapnik}%
{A presentation of the two major FOSS alternatives: Mapserver and Geoserver}%
{Sven Geggus is long-time GIS and open source evangelist.}%
{Talking about Map rendering and tileservers in an environment of Openstreetmap activists makes it look like if there is only one piece of software which can be used to do this job: Mapnik.

This talk will focus on the two major FOSS alternatives, which are less well
known outside the FOSSGIS/OSGeo community: Geoserver and Mapserver

I will show some simple rendering examples for both of them (and Mapnik as a reference). Finally I will present advantages and disadvantages of each particular software.}


\talkabstract{The use of OpenStreetMap data in journey planning}%
{Example on the usability of the data for railwayrouting}%
{Thomas Jakubicka is responsible for project management at Mentz Datenverarbeitung. He is especially involved in many projects related to OSM.}%
{The continually improving accuracy, grade of detail and completeness in OSM data on the one hand and the always rising demand for data with higher information density, has drawn the attention of public transport operators and autorities towards OSM. Furthermore open governmental data becomes more and more important to public transport services and the idea of OSM supplements this approach.
Mentz Datenverarbeitung (mdv) is developing journey planning systems and therefore has recently implemented the support of OSM data into their systems. A key aspect of journey planning is the georeferencing of the data in a GIS. In this talk the various aspects of railway and tramway routing will be presented. The intention is, to take the example of rail-bound routing and discuss the challenges as well as the benefits of OSM data for public transport and the implications for the different stakeholders: public transport providers, public transport users and the broad OpenStreetMap community.}


\talkabstract{pgmapcss - Advanced cartography for Mapnik}%
{pgmapcss implements MapCSS for Mapnik - by moving the cartography into the database.}%
{}%
{pgmapcss combines MapCSS - a more or less standardized map description language - with Mapnik - a widely used map renderer for vector or bitmap maps. In contrast to other attempts at using CSS-like styling, the actual cartography process (evaluating which map features should be rendered  in the current view and how) is moved into a database function (Using PL/Python3 in a PostgreSQL database). Mapnik just needs to read the final properties - e.g. geometry, widths, colors, texts, …. This database function is compiled by pgmapcss from the MapCSS style sheet.

\noindent There are many advantages to this idea:
\begin{itemize}
\item Writing MapCSS simplifies writing style sheets for Mapnik, as there's no longer a separation into database query and styling. Often those database queries were complex and required advanced knowledge of SQL. pgmapcss takes most of the complexity away. In addition, the actual database queries for map features are optimized for each zoom level.
\item MapCSS is powerful because map features can be related to each other (e.g. by relation membership or the proximity of objects). Furthermore calculations are possible (e.g. show the width of a line based on the value of a tag of the map feature). pgmapcss even offers geometric calculations (e.g. create lines between map features, rotation, buffers, …).
\item As there are more than ten libraries which implement MapCSS, the language is available on different platforms and for different output media. Styles can be used with various libraries,  although there are differences between dialects, but hopefully they will converge in the future.
\end{itemize}

Homepage: \url{https://github.com/plepe/pgmapcss}}


\talkabstract{State of the Kort Game}%
{The First OpenStreetMap Mobile Mini Game Goes Public}%
{The author is professor for information systems at the University for Applied Sciences (UAS/HSR) in Rapperswil (Switzerland). He leads the (OSGeo) Geometa Lab at Institute for Software. He teaches database management systens as well as geographic information systems (GIS) at Bachelor and Master level, as well as in advanced studies. In his projects often Open Source and volunteered geographic information (VGI, crowdsourcing) play a role - but not only. }%
{Kort ist a mobile web app to fix OpenStreetMap data. It runs on the most commond browsers. The app uses the concept of gamification. Game-like elements like points (so called 'Koins') are collected by the players by fulfilling a mission, like adding names to POIs without one. All proposed solutions are validated by other players. If 3 players aggreed on a proposal it is integrated on OpenStreetMap.

This is a report about our efforts to let volunteers contribute additional mission types. Besides restructions of mission types and internal refactoring issues, the main question is, how to design the user and machine-oriented interface. (http://www.kort.ch)}


\talkabstract{Lightning Map Tiles}%
{A Vector Approach to Raster Maps}%
{Andy Allan created the OpenCycleMap map layer in 2007, and is now providing high-quality map tiles to hundreds of applications and websites through his Thunderforest mapping platform. }%
{Many of the maps you see created from OpenStreetMap data use the battle-hardened mod\_tile software stack to render images on-demand directly from a PostGIS database. To work around some of its inherent limitations, Andy started working on an alternative  approach based on the mapnik-vector-tile software, storing the map data in protocol-buffer based vector tiles. This opens up a range of cartographic features, from faster map drawing times to easing the burden of hosting multiple styles. Andy will discuss the approaches taken, the pitfalls encountered and the future possibilities made possible using this vector tile approach.}


\talkabstract{OSM housenumber evaluation}%
{how does it help to improve housenumber quantity and quality}%
{Dietmar Seiffert is mapping OSM for about 4 years, active in forum and providing OSM streetlist and housenumber evaluations.}%
{housenumbers will be inserted more and more in OSM. This tool gives you a feedback which housenumber are still missing and allows to give feedback to the administration of the town. You just need to ask for a housenumber list at your village or town administration}

\talkabstract{Rendering Maps with OpenGL}%
{How to render a vast dataset like OSM on mobile devices}%
{After studying IT-Systems Engineering in Potsdam, Konstantin Käfer joined Mapbox and works on map rendering}%
{Rendering maps on the client becomes more important as technologies emerge and devices get faster. While most maps are still served as raster tiles, there is a case for delivering vectors to the client. In this talk, I'll review existing rendering techniques and data formats, as well as point out their advantages and disadvantages.}

\talkabstract{State of the OSM-France's tools}%
{OSM-France develop and maintain tools, let's take a tour}%
{Christian Quest, 48, lives near Paris. He has been an OSM contributor
since 2009, and he's a founding member and the current president of
OSM-FR. Focused on data quality and exhaustivity, working on several
renderings where external data are compared to OSM data to find "white
spots", missing data as well as OSM-FR styled tiles. Christian is also
part of OSM-FR tech team as system administrator taking care of a dozen
of servers. He discovered OSM through his other (past) activities:
paragliding and genealogy.}%
{Some tool available on french servers are generic like taginfo, tiles server or overpass API. But some are specific, but useful for all. Among other we can find HOT tile server, uMap, API proxy, cadastre extractor, polygon generator, extract of new area with diff, boundary maker, Osmose \dots}



\talkabstract{Tesselator’s Delight}%
{Introduction to OpenStreetMap for WebGL}%
{Brett Camper is interested in graphics programming, data visualization, game and interface design, and related areas. He recently helped start Mapzen, a company focused on developing tools and apps to improve open source geo.}%
{In the past couple years, WebGL has moved from an emerging technology to a widely supported one, providing a promising and powerful toolkit for rendering OSM data in the browser. But WebGL is often daunting and foreign to web programmers, and surprisingly few resources are available to learn. In this talk I hope to help demystify WebGL for OSM, introducing the basics of a rendering pipeline using open source code examples, including: getting data from the server via GeoJSON or binary vector tiles, turning OSM geometries into WebGL-compliant triangle primitives ("tessellation"), constructing neatly joined polygonal line segments for roads and other pathways, extruding building outlines into 3D models, and using WebGL vertex and fragment shaders to alter effects like lighting direction and viewer perspective. All presented code is open source and also available as a plugin for Leaflet JS, making it easily accessible to developers who want to experiment with adding WebGL and 3D components to their maps.}


\talkabstract{Osmose}%
{A QA tools for detect and fix errors and integrate OpenData}%
{Frédéric Rodrigo is secretary of OpenStreetMap France, freelance in geomatic, works on quality tools and development project around OSM.}%
{Osmose is one of many quality assurance tools available to detect errors and inconsistencies in OpenStreetMap data. It is also useful for integrating OpenData. Osmose has more than 250 checks. Number of analysers and countries supported is is still rising. With last funding Osmose have an integrated tags editor usable on desktop and mobile.}

\talkabstract{OSM Buildings}%
{Now and next}%
{Jan Marsch is from Berlin, Germany and has been software engineer for desktop and mobile web applications for about 16 years. By doing some projects for Nokia Maps five years ago, he got addicted to maps. His technical focus is on JavaScript, HTML5, Canvas, REST, PHP and databases. Other strenghts are entrepreneural thinking, performance optimization and user experience. He is happy working with small companies but also did long term projects for Here, Daimler, Bayer, Deutsche Lufthansa and TNT.}%
{I'll give you an introduction to OSM Buildings, a project to visualize OpenStreetMap building geometry in modern web browsers.
You'll learn what its render modes are and how easily it integrates with existing web map engines.

And if you are familiar with the project already, there are a lot new data sources to discover and user interaction got a big step forward.

Furthermore I’ll explain how OSM Buildings compares to similar projects and where it stands out. For future plans, you'll see why there won't be another gray 3D blocks engine.

Meanwhile, follow @OSMBuildings on Twitter!}


\talkabstract{MapRoulette 2: Electric Boogaloo}%
{It's back, it's real, and it's fun}%
{Serge Wroclawski is a longstanding member of the OpenStreetMap community. He's one of the founders of OSM US, a DWG member. He's also the author of "Why the World Needs OpenStreetMap".}%
{MapRoulette has been undergoing a major rewrite for nearly two years, but the new version is finally out. In this presentation, you will learn about MapRoulette's history and about what's new in the current version, as well as what the future may hold.}

\talkabstract{2.5D maps and bird views with OpenStreetMap and Blender}%
{How to create 2.5D maps and bird views with OpenStreetMap data and Blender open source software}%
{Vladimir Elistratov works as a developer of web mapping applications. He is OSM contributor since 2008. }%
{2.5D maps are ordinary 2D maps in the web Mercator projection enhanced with a layer of 3D buildings rendered in the oblique projection. A birdview is an attractive way to realistically represent neighborhood. Realistic 3D models of buildings are used in both cases.

Blender open source software is used for both 3D modelling and 3D rendering. To perform georeferencing of each 3D building, OSM data are imported into Blender as flat objects in the transverse Mercator map projection. One places a 3D model at the correct position by moving and rotating it over the imported OSM data. The corresponding geographic coordinates and rotation angle are saved in a Blender file or in a database. Whole cities can be composed in this way.

There are two methods to create 2.5D maps. In the first method every 3D object is rendered in the oblique projection as a separate graphic symbol. Mapnik's PointSymbolizer is employed to render those images at the center of the corresponding object. In the second method all 3D objects are rendered in the oblique projection in one image. The image is georeferenced and rendered on the map as a raster layer. In both methods 3D objects are scaled in accordance with their latitude to take distortion of the Mercator projection into account.

3D objects must be placed on the terrain for bird views. SRTM data are used for that purpose. If necessary, terrain data are edited manually in Blender. Another key problem for bird views is to develop attractive 3D representation for numerous street objects like trees, street lights, fences, benches, bus stops, etc.}

\talkabstract{}%
{}%
{}%
{}

\talkabstract{An Approach for Cartographically Plausible and Good Point-Feature Label Placement on OSM-based Maps}%
{A multi-criteria model with practical implementation}%
{Maxim Rylov is a PhD student in GIScience Research Group, Department of Geography, Heidelberg University. His main research interests are digital and web cartography, automated label placement, computational geometry and GIS mapping.}%
{The cartographic label placement is a very important aspect of a map production. This task is essential for both traditional and automated cartography. Every OSM-based map is annotated using a labeling algorithm in one of the existing toolkits for rendering maps.  However, none of these algorithms is able to take into account a rich set of well-established cartographic guidelines for feature annotation used by human cartographers. Assigning names to point-features is one of the map lettering tasks. In our talk we present an approach, expressed as a multi-
criteria optimization model, that complies with almost all well-defined cartographic placement principles and requirements for point-feature label placement. We show how this approach allows significantly increase toponym density on an OSM-based map without effecting readability and legibility.}


